annals_manuscript

### 2020-04-03 annals_manuscript.Rmd
updated

# Alpha and beta diversity figure
```{r}

@@

# gets the meta data 
  pseq.meta <- meta(pseq.subset)
  # generates the diversites
  tab <- microbiome::alpha(pseq.subset, index = "shannon")

  # This country versus nationality
  p3<-   ggviolin(pseq.meta, x = "nationality", y = "Shannon",
                  add = "boxplot", fill = "nationality", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")

  p4 <<- ggviolin(data=subset(pseq.meta, !is.na(mnd)), x = "mnd", y = "Shannon", add = "boxplot", fill = "mnd", na.rm = TRUE, palette = c("yellow", "orange", "red"))

  p4 <<- p4 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND DALYs") + theme(legend.position="none")

p.1 <<-plot_grid(title, p1, p3, p2, p4, align = "v" , ncol = 2, labels = c('','A', 'B', 'C', 'D'), rel_widths = 1, rel_heights = 2)
print(p.1)

@@

## BETA DIVERSITY
Beta-diversity: Measures for differences between samples from different groups to identify if there are differences in the overall community composition and structure.
WE want to: quantify community divergence within a given sample set
Divergence of a given sample can be quantified as the average dissimilarity of each sample from the group mean; the dissimilarity can be quantified by beta diversity.

>>>>

# Alpha figure
```{r}

pseq.meta <- meta(pseq.subset)
  # generates the diversites
  tab <- microbiome::alpha(pseq.subset, index = "shannon")

  print("Second graph has completed")

  p3 <<-plot_grid(p1,p2, align = "v" , ncol = 1, axis = 'tblr', labels = c('C', 'E'),rel_heights = c(5,5,5))

print(p3)

@@

p4 <<-plot_grid(p1,p2, align = "v" , ncol = 1, axis = 'tblr', labels = c('D', 'F'),rel_heights = c(5,5,5))

print(p4)

p6 <<-plot_grid(p3,p4, align = "v" , ncol = 2, rel_heights = c(10,10,10))

print(p6)

# Composition and Alpha 
```{r}

p <- plot.p.new.rel + scale_fill_manual(values=c( "#A5CEE3", "#1F78B4", "#34A02C", "#E3211D", "#FF7E00", "blue", "yellow", "red"))

p1 <- plot.p.new.rel + scale_fill_brewer(palette = "Paired")

p7 <<-plot_grid(p,p1, align = "v" , ncol = 2, labels = c('A', 'B'),rel_heights = c(10,10,10))
plot(p7)

p8 <<-plot_grid(p7, p6, align = "v" , nrow = 2, rel_heights = c(10,30,30))

print(p8)

#Beta
```{r}

 p <<-  boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))
 # p1 = plot_ordination(pseq.subset, pseq.nmds, type = "taxa", color = "mnd", title = "Taxa by Phylum") # I dont think that we need this any more. 
  #p1 + facet_wrap(~Phylum, 3)
  pseq.nmds <- ordinate(pseq.subset, "NMDS", "bray")
  p2 <- plot_ordination(pseq.subset, pseq.nmds, type = "sample", color = "mnd")

   #PCoA for compositional data with Bray-Curtis distances
    #   p4 <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", colour = "mnd", size = 3) 
    #   p4 <- p4 + scale_color_brewer(palette = "Dark2")+ scale_fill_gradient(low = "#e0ecf4", high = "#6e016b") 
    #p6 <- plot_landscape(pseq.subset, method = "MDS", distance = "bray", col = "mnd", size = 3) 
  #   # PCoA for compositional data with Bray-Curtis distances
   #p6 <- plot_landscape(microbiome::transform(pseq.core, "compositional"),
    #                    method = "PCoA", distance = "bray") +
    #     labs(title = paste("PCoA / Compositional / Bray-Curtis"))
  # 
    tsne_res <- tsne_phyloseq(pseq.subset, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
    p7 <- plot_tsne_phyloseq(pseq.subset, tsne_res,color = 'mnd')
    p8 <- p7 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "TSNE",color = "MND DALYs") + theme(legend.position = "bottom")
    save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), p8)

    p10 <<-plot_grid(title,p3, p5,p8, ncol = 1, labels = c('','A', 'B','C'),rel_heights = c(0.3, 1.5,1.5,2))

    print(p10)

# Permanova figure
```{r}

 p.low <<- subset_samples(pseq.subset, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(pseq.subset, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(pseq.subset, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))

  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 12)  + annotate("text", x = 2.5, y = 0.01, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa") + scale_fill_discrete(name = "Taxa") + theme_classic() + theme(axis.text.y = element_text(face="italic"))

plot(p, xlim=c(-0.03,0.03))

  # scale_x_continuous(limits=c(-0.03, 0.03), breaks=c(-0.03, -0.02, -0.01, 0, 0.01, 0.02, 0.03))

 p.low <<- subset_samples(atlas.pseq, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(atlas.pseq, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(atlas.pseq, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))

  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p2 <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 12)  + annotate("text", x = 2.5, y = 0.01, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa") + scale_fill_discrete(name = "Taxa") + theme_classic() + theme(axis.text.y = element_text(face="italic"))

perm <- plot_grid(p,p2, align = "v" , ncol = 1, axis = 'tblr', labels = c('A', 'B'),rel_heights = c(10,10,10))

print(perm)
ggsave("permanova_annals.png", device = "png",width = 250, height = 280, units = "mm")

@@

BETA DIVERSITY
Beta-diversity: Measures for differences between samples from different groups to identify if there are differences in the overall community composition and structure.
WE want to: quantify community divergence within a given sample set
Divergence of a given sample can be quantified as the average dissimilarity of each sample from the group mean; the dissimilarity can be quantified by beta diversity.
